---
output: 
  pdf_document:
    latex_engine: xelatex
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Helvetica}
---


# Platform 2: R

In R, the rpart package is used to create the Decision Tree model, and rpart.plot is used for visualizing the tree structure. The Gini criterion, which measures the impurity of splits, is used in the Decision Tree model by default in R.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(rpart)
library(rpart.plot)
library(caret)
library(ggplot2)
```

## Data Loading

The cleaned dataset from the Jupyter Notebook is used to maintain consistency across both platforms.

```{r}
data <- read.csv("cleaned_data.csv")
```

## Data Preprocessing
The dataset is split into training and testing sets to allow for model evaluation. 70% of the data is used for training, while 30% is reserved for testing. The set.seed(42) ensures reproducibility.
```{r}
# Splitting the dataset
set.seed(42)
trainingIndex <- createDataPartition(data$HadHeartAttack, p = 0.7, list = FALSE)
train <- data[trainingIndex, ]
test <- data[-trainingIndex, ]
```

## Model Building
A Decision Tree is built using the rpart function with a maximum depth of 5, consistent with the Python model, and the Gini criterion (default in R). The rpart.plot function is used to visualize the tree, showing significant risk factors for heart disease.
```{r}
dt_model <- rpart(HadHeartAttack ~ ., 
                  data = train, 
                  method = "class", 
                  control = rpart.control(maxdepth = 5, cp = 0.001, minsplit = 20))

# Print summary of the model to check variable importance and splits
summary(dt_model)

```

## Model Evaluation
The model is used to predict the outcomes in the test set, and a confusion matrix is generated to evaluate the accuracy of the model, comparing predicted and actual outcomes。
```{r}
# Make predictions
predictions <- predict(dt_model, newdata = test, type = "class")

# Convert both to factors and ensure they have the same levels
predictions <- as.factor(predictions)
test$HadHeartAttack <- as.factor(test$HadHeartAttack)

# Ensure they have the same levels
levels(predictions) <- levels(test$HadHeartAttack)

# Evaluate the model
confusionMatrix(predictions, test$HadHeartAttack)
```
## Model Visualization
The rpart.plot function is used to visualize the tree, showing significant risk factors for heart dise
```{r}

# Plot the decision tree with enhanced visualization
rpart.plot(dt_model, main = "Decision Tree for Heart Disease Risk Factors", extra = 106)
```

```{r}
# Extract variable importance
importance <- as.data.frame(dt_model$variable.importance)
# Convert row names to a proper column for plotting
importance$Variable <- rownames(importance)
# Rename the columns for clarity
colnames(importance) <- c("Importance", "Variable")
# Sort the factors by importance in descending order
importance <- importance[order(-importance$Importance), ]
# Select the top 10 most important factors
top_10_importance <- importance[1:10, ]
# Plot using ggplot2
ggplot(top_10_importance, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip the coordinates to have a horizontal bar plot
  theme_minimal() +
  labs(title = "R - Top 10 Significant Risk Factors", x = "Risk Factors", y = "Importance")
```